---
title: "case studies info is beautifull"
author: "Tomasz Olczyk"
date: "3/30/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Studia przypadku z info is beaudifull


### Hollywood hits and flops

https://informationisbeautiful.net/visualizations/hollywood-2023-hits-flops/


położenia etykiet, skale kolorów, kształy linie

adnotacje w postaci kształtów

dane: 

https://docs.google.com/spreadsheets/d/12bnGB7w5T03f7Bq1PG6gqFIdsLccdP6E-86xdVjbCOc/edit#gid=423546026


### most common passwords

https://informationisbeautiful.net/visualizations/top-500-passwords-visualized/

wordloud na jajku według kolorów dość bazowe ale ciekawe wizualnie można dodać jako zadanie zaawansowane
ew. coś na zajęcia



### prawda fałsz w filmach

dane, skala kolorów, geom segment geom segment informacyjnie, czas liczony w minutach sekundach i jego przeliczanie

https://docs.google.com/spreadsheets/d/1sJDpzYH_sMYuYHqkmZeJGIq_TEXGDjboYdSoew7UjZ8/edit#gid=1939779101


### ukraina w liczbach

kolory, kombinacje wykresów, treemaps może na zadanie końcowe?

https://informationisbeautiful.net/visualizations/ukraine-russian-war-infographics-data-visuals/#two-years


https://docs.google.com/spreadsheets/d/17YhLXD459Q8_8ez0aAl9EP4EbwgYndTDeMgJzvp5G_w/edit#gid=445996165



### analiza sieciowa middle east key players noable  relationsips

Trochę zaawansowanej wizualizacji : wymaga kontrolowania typu więzi  i typu węzła
https://informationisbeautiful.net/visualizations/the-middle-east-key-players-notable-relationships/

https://docs.google.com/spreadsheets/d/1cu0V71960pnrkf5meIavnsoKJcvCRQTV8Y-koODrPc8/edit#gid=8


nad danymi trzeba by popracować


### mapa, który kraj jest w czym najlepszy


zagadka jak dodać etykietę do kraju z linią może ggrepel, a może wystarczy 
https://docs.google.com/spreadsheets/d/11uifsxtHKwRysrxNxTDhvWLDHTlxQ0jYP8PODoLM2hM/edit#gid=0


### mapa  energia odnawialna

https://informationisbeautiful.net/visualizations/renewable-energy-and-electricity/

Dane:
można wykorzystać do ćwiczenia z mapami typu kartogram:

https://docs.google.com/spreadsheets/d/1j8E3rPujJHWgrLCX8ushGQizIlxPn1v4YxPyMoMgMXY/edit#gid=2077503762


### geometrie, osie kolory

https://informationisbeautiful.net/visualizations/mountains-out-of-molehills/


### Circular voronoi

https://informationisbeautiful.net/visualizations/ukraine-russian-war-infographics-data-visuals/#OilReserves


```{r}
install.packages("voronoiTreemap")
```


```{r}
library(voronoiTreemap)

vor <- data.frame(h1 = 'World', 
                  h2 = c('Europe', 'Europe', 'Americas', 'Asia',
                         'Europe', 'Other'),
                  h3 = df$country,
                  color = hcl.colors(nrow(df), palette = 'TealRose'),
                  weight = df$prod,
                  codes = df$country)

vt <- vt_input_from_df(vor)

vt_d3(vt_export_json(vt))
```


# arcdiagram

podróże w czasie



https://docs.google.com/spreadsheets/d/1XcwuvLzlq3KEauW7LnbFGC2oleHR5TJOmSYhhsJwv5k/edit?hl=en_GB&pli=1&hl=en_GB&pli=1#gid=0


https://informationisbeautiful.net/2009/time-travel/



It's relatively easy to make a Voronoi tasselation in R, but it's harder to make a Voronoi treemap. The linked Q&A does it by using the voronoiTreemap package, which is essentially just a wrapper round a JavaScript library. As far as I can tell, this is the only published R package that generates Voronoi treemaps.

Our two options are to calculate the polygons ourselves from scratch, or somehow extract the polygons from the SVG output of voronoiTreemap.

With regards to the first option, this is not a trivial problem. To see just how complex it is, and also to get a fully worked solution in R, you can check out this fantastic article by Paul Murrell. The code runs to several pages and is over a decade old, so I'm not sure if all the dependencies still work. It's disappointing that no-one has put it all together in a package on CRAN, but perhaps it's a bit niche.

If you struggle with Paul Murrell's approach, you are left trying to harvest polygons from the output of voronoiTreemap. Although this package works well, the output does not lend itself to being harvested for polygons, and we do not get access to the intermediate calculations that would allow us to generate the polygons ourselves in R. It's not impossible, and there are a few ways to tackle it, but they are all fairly convoluted.

The following approach starts with plotting your treemap as normal with voronoiTreemap, but without the labels:

```{r}
library(voronoiTreemap)
library(terra)
library(tidyverse)

```

```{r}

df <- data.frame(country = c("Ukraine", "Russia", "Argentina", 
                             "China", "Romania", "Other"),
                 prod = c(11.0, 10.6, 3.1, 2.4, 2.1, 15.3))

vor <- data.frame(h1 = 'World', 
                  h2 = c('Europe', 'Europe', 'Americas', 'Asia',
                         'Europe', 'Other'),
                  h3 = df$country,
                  color = hcl.colors(nrow(df), palette = 'TealRose'),
                  weight = df$prod,
                  codes = "")

vt <- vt_input_from_df(vor)

v <- vt_d3(vt_export_json(vt))

v
```





