---
title: "gender pay gap uk"
author: "Tomasz Olczyk"
date: "3/30/2024"
output: html_document
params:
  category:
    value: all
    choices: [all, admin & organisation, 
    care & education, creative & media, law & justice,
     manual work, sales & serving others, science, tech & engineering,
     senior managers & execs]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


```{r}
pay_gap <- read_csv("https://raw.githubusercontent.com/Tomasz-Olczyk/wizualizacjaR/main/case%20studies%20/pay_gap_uk.csv")
head(pay_gap)
```



```{r}
dane_gap <- pay_gap %>%
  select(occupation, 
        women = women_average_annual_salary, 
         men = men_average_annual_salary,
        pay_gap,
        category) #%>%
 # mutate(category = factor(category))
```


```{r}
dane_long <- pay_gap %>%
  select(occupation, 
         women = women_average_annual_salary, 
         men =men_average_annual_salary, 
         pay_gap,
         category)  %>%
  pivot_longer(cols = c(women, men), 
               names_to  = "gender", 
               values_to = "salary")# %>%
  #mutate(category = factor(category))

```

```{r}
head(dane_long)
```


kolory <-
```{r}
kolor1 <- rgb(80, 27, 60, maxColorValue = 255)
kolor2 <- rgb(186, 195, 106, maxColorValue = 255)

skala <- c(kolor1, kolor2)
```



## wykres według pensji

```{r}

x <- 
ggplot(data = dane_long %>% 
         filter(category == "admin & organisation"), aes(y=reorder(occupation, salary))) +
  geom_segment(data = dane_gap %>% 
         filter(category == "admin & organisation"), 
         aes(y = occupation, 
                                    yend=occupation, 
                                    x= women, 
                                    xend = men)) +
  geom_point(aes(x = salary, colour = gender)) +
  #scale_color_manual(values = skala, guide_legend(position = "top")) +
  theme_void() 
  
```



```{r}
x
```


# wykres według pay gap

```{r}
y <- dane_long %>% 
  filter(category == params$category) %>% 
  ggplot( aes(y=reorder(occupation, pay_gap))) +
  geom_segment(data = dane_gap, aes(y = occupation, 
                                    yend=occupation, 
                                    x= women, 
                                    xend = men)) +
  geom_point(aes(x = salary, colour = gender)) 
  
```


```{r}
y
```


```{r}
library(plotly)
```



```{r}

ggplotly(y)

```

