---
title: "studia przypadków"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Wykres Benjamina Nowaka

Inforgrafika nawiązująca z Panteonu nawiązujący do jednego z [wykresów](https://bjnnowak.github.io/memo_agro/) Benjamina Nowaka dotyczących produkcji rolnej.

### Biblioteki

```{r}

library(tidyverse)


library(waffle) # biblioteka do wykresów typu waffle
library(sf) # mapy
library(ggtext)

```


biblioteka do przyporządkowania krajów. W przypadku danych panteon poniższa biblioteka nie jest potrzebna bo kontytnenty są już przypisane do krajów więc można narysować mapę bez dodakowych złączeń


```{r}
install.packages("rnaturalearth")
library(rnaturalearth)
```

### dane Nowaka


```{r}
data<-read_csv('https://raw.githubusercontent.com/BjnNowak/book/main/data/wheat_maize_2010_2020.csv')
```

```{r}
clean<-data %>%
  # Changing column name to avoid space
  dplyr::rename(M49 = 'Area Code (M49)')%>%
  drop_na(Value)%>%
  group_by(M49, Item)%>%
  # Mean yield in tons
  summarize(yield=mean(Value)/10000)%>%
  ungroup()

head(clean)
```


```{r}
wrld <- ne_countries(type = "countries", scale = "small", returnclass = "sf")%>%
  # Drop geometry and keep only attribute table
  st_drop_geometry()

clean<-clean%>%
  # Join by UN code
  left_join(wrld,by=c('M49'='un_a3'))%>%
  select(M49,Item,yield,continent)%>%
  drop_na(continent)

head(clean)
```


```{r}
clean<-clean %>%
  mutate(
    rd=floor(yield),
    ct=1) %>%
  group_by(rd, Item, continent) %>%
  summarize(n=sum(ct)) %>%
  ungroup()

head(clean)
```



```{r}
ggplot(
    # Keep only data for wheat
    clean%>%filter(Item=="Wheat"),
    aes(values=n, fill=continent)) +
  waffle::geom_waffle(n_rows = 4,              # Number of squares in each row
    color = "white",         # Border color
    flip = TRUE, na.rm=TRUE) +
  facet_grid(~rd)+
  coord_equal()
```


```{r}
pal_fill <- c(
  "Africa" = "#FFC43D", "Asia" = "#F0426B", "Europe" = "#5A4FCF",
  "South America" = "#06D6A0", "North America" = "#059E75",
  "Oceania" = "#F68EA6",
  # Set alpha to 0 to hide 'z'
  'z'=alpha('white',0)
)

pal_color <- c(
  "Africa" = "white", "Asia" = "white", "Europe" = "white",
  "South America" = "white", "North America" = "white",
  "Oceania" = "white",
  # Set alpha to 0 to hide 'z'
  'z'=alpha('white',0)
)
```


```{r}
cpl<-clean%>%
  # Removing some countries with unrealistic yield for maize
  filter(rd<14)%>%
  # Complete all combinations for real continents, but with 0 value 
  # (this prevents cases from being created during the 2nd application of the function)
  complete(
    rd,Item,continent, 
    fill=list(n=0)
  )%>%
  # Add imaginary 'z' continent
  add_row(continent='z',Item='Wheat',n=0,rd=1)%>%
  # Complete all combination for the 'z' continente
  complete(
    rd,Item,continent, 
    fill=list(n=1)
  )
```




```{r}
cpl$Item <- as.factor(cpl$Item)
cpl$Item <- fct_relevel(cpl$Item, "Wheat","Maize (corn)")
```


```{r}
p1<-ggplot(
  cpl,
  aes(values=n,fill=continent,color=continent))+
  waffle::geom_waffle(
    n_rows = 3, 
    flip = TRUE, na.rm=TRUE)+
  facet_grid(
    Item~rd, 
    # Switch both facet position
    switch="both"
  )+
  scale_x_discrete()+
  scale_fill_manual(values=pal_fill)+
  scale_color_manual(values=pal_color)+
  coord_equal()

p1
```

```{r}
title<-"<b>The Yield Gap</b>"
sub<-"Each square is a country, colored according to continent:<br><b><span style='color:#FFC43D;'>Africa</span></b>, <b><span style='color:#F0426B;'>Asia</span></b>, <b><span style='color:#5A4FCF;'>Europe</span></b>, <b><span style='color:#059E75;'>Central and North America</span></b>, <b><span style='color:#06D6A0;'>South America</span></b> or <b><span style='color:#F68EA6;'>Oceania</span>. "
```

```{r}
p2<-p1+
  # Hide legend
  guides(fill='none',color='none')+
  # Add title and subtitle
  labs(title=title,subtitle=sub)+
  theme(
    # Enable markdown for title and subtitle
    plot.title=element_markdown(),
    plot.subtitle=element_markdown(),
    # "Clean" facets 
    panel.background=element_rect(fill="white"),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    strip.background.x = element_rect(fill="white"),
    strip.background.y = element_rect(fill="dimgrey"),
    strip.text.y = element_text(color="white"))

p2
```


### dane Panteon

```{r}

panteon <- read_csv("https://raw.githubusercontent.com/Tomasz-Olczyk/wizualizacjaR/main/podstawy/panteon_s.csv")

```


#### analogiczne przeliczenia do PAnteona



```{r}
czyste <- panteon %>%
  group_by(countryCode3, gender)%>%
  summarize(wyświetlenia=mean(AverageViews)) %>%
  ungroup()

```


```{r}
kontynenty <- panteon %>%
  select(countryCode3, continentName)
```



```{r}
pan <- pant %>%
  mutate(
    rd=floor(wyświetlenia),
    ct = 1) %>%
  group_by(rd, gender, continentName) %>%
  summarize(n=sum(ct)) %>%
  ungroup()
```


```{r}

pant <-  left_join(czyste, kontynenty, by = "countryCode3")
```

```{r}
p3<-ggplot(pan,
  aes(values=n,fill=continentName,color=continentName))+
  waffle::geom_waffle(
    n_rows = 3, 
    flip = TRUE, na.rm=TRUE)+
  facet_wrap(gender~rd) 
    # Switch both facet position
    
p3
```

